---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(g2r)
g2r_globals(font = "Quicksand")
```

## Synthax

The synthax of g2r is heavily inspired by ggplot2.

```{r, echo=FALSE}
df <- dplyr::tribble(
  ~"ggplot2",       ~"g2r",
  "ggplot",         "g2",
  "aes",            "asp",
  "scale_*",        "gauge_*",
  "geom_*",         "fig_*",
  "facet_*",        "plane_*",
  "theme_*",        "motif_*"
)

knitr::kable(df, caption = "ggplot2 to g2r synthax map.", format = "html")
```

## Aspects

In g2r you use aspects (`asp`) instead of aesthetics (`aes`) in ggplot2 but they work in very much the same way.

```{r}
g2(cars, asp(dist, speed, size = speed)) %>% 
  fig_point()
```

Then you can use gauges (instead of scales in ggplot2) to manipulate those aspects.

```{r}
g2(cars, asp(dist, speed, color = speed, size = 10)) %>% 
  fig_point() %>% 
  gauge_color(c("blue", "white", "red"))
```

It may seem like some some figures (equivalent to geoms) are missing, but technically all are available.

```{r}
fruits %>% 
  g2(asp(value, color = fruit)) %>% 
  fig_interval_stack() %>% 
  coord_type("theta")
```

You won't find `fig_bar` but `fig_interval` does the trick. Below we the `adjust` function which is similar to using the `stat` argument in ggplot2.

```{r}
g2(temp, asp(month, temp, color = city)) %>% 
  fig_interval_dodge() 
```

There are no fig_boxplot either, yet.

```{r, collapse=TRUE}
library(dplyr)
library(purrr)

df <- rep(100, 8) %>% 
  map(runif) %>% 
  map(boxplot, plot = FALSE) %>% 
  map_dfr(function(x){
    x$stats %>% 
      unlist() %>% 
      t %>% 
      as.data.frame
  }) %>% 
  mutate(
    group = LETTERS[1:n()],
    type = rep(letters[1:2], 4)
  ) %>% 
  tidyr::nest(V1, V2, V3, V4, V5)

df$data <- df$data %>% 
  map(unlist) %>% 
  map(unname)
```

Plot created data

```{r}
g2(df, asp(type, data, adjust = "dodge", shape = "box", size = 60, color = group)) %>% 
  fig_schema()
```

There is a `fig_violin` but it expects the data in a specific format.

```{r}
get_density <- function(k = "A"){
  density <- runif(100) %>% 
    density() 
  df <- dplyr::tibble(key = k)
  df$x <- list(density$x)
  df$y <- list(density$y)
  return(df)
}

df <- get_density("A") %>% 
  bind_rows(get_density("B")) %>% 
  bind_rows(get_density("C"))

g2(df, asp(key, x, size = y, color = key)) %>% 
  fig_violin(style(fillOpacity = .3)) %>% 
  conf_tooltip(FALSE)
```

A heatmap is easier.

```{r}
library(dplyr)

data("diamonds", package = "ggplot2")

palette <- c("blue", "cyan", "lime", "yellow", "red")

diamonds %>% 
  count(carat, price) %>% 
  g2(asp(carat, price, color = n)) %>% 
  fig_heatmap() %>% 
  gauge_color(palette)
```

```{r}
iris %>%
  g2(asp(Petal.Length, Petal.Width, color = Species)) %>% 
  fig_point() %>%
  plane_wrap(planes(Species), type = "tree")
```

If you navigate the [original examples of the g2 JavaScript](https://antv.alipay.com/zh-cn/g2/3.x/demo/index.html) you will notice that they often make use of the [dataset library](https://github.com/antvis/data-set) to transform the data. This is imprelement as the function `alter`, mainly to easily replicate examples. Note that the waffle chart below can now simply be done with `fig_waffle`.

```{r}
altered_fruits <- fruits %>% 
  mutate(value = value * 100) %>% 
  select(name = fruit, value) %>% 
  alter(type = "waffle", maxCount = 500, rows = 12) 
 
g2(altered_fruits, asp(x, y, shape = "square", color = name, size = 10)) %>% 
  fig_point(axes = FALSE) %>% 
  gauge_axis(x, FALSE) %>% 
  motif(padding = rep(50, 4)) 
```

The `alter` function is used internally by `get_map` to process GeoJSON files. It'll enable you to easily build otherwise difficult charts. Though there is a `fig_hex` helper now.

```{r}
g2(gaus, asp(x, y)) %>% 
  fig_bin() %>% 
  gauge_color("#BAE7FF-#1890FF-#0050B3")
```

```{r}
df <- dplyr::tibble(
  x = runif(25, 1, 500),
  y = runif(25, 1, 500),
  value = runif(25, 1, 500)
)

g2(df, asp(x, y, color = value)) %>%
  fig_voronoi(axes = FALSE)
```

```{r}
fruits %>% 
  arrange(value) %>% 
  mutate(label = fruit) %>%  
  g2(asp(fruit, value)) %>% 
  fig_interval(asp(shape = "line")) %>%
  fig_point(asp(shape = "circle")) %>%  
  coord_type("theta", innerRadius = .2, endAngle = pi) %>% 
  info_text(asp(fruit, 0, content = label), style = list(textAlign = "right"))
```

Many figures have been implemented, many of them actually use the `group` aspect. 

```{r}
g2(iris, asp(Sepal.Length, group = Species, color = Species))  %>%
  fig_histogram(bin_width = .3)
```

```{r}
g2(iris, asp(Petal.Length, group = Species, color = Species)) %>%
  fig_density()
```

```{r}
df <- data.frame(
  trt = factor(c(1, 1, 2, 2)),
  resp = c(1, 5, 3, 4),
  group = factor(c(1, 2, 1, 2)),
  upper = c(1.1, 5.3, 3.3, 4.2),
  lower = c(0.8, 4.6, 2.4, 3.6)
)

g2(df, asp(trt, resp, ymin = lower, ymax = upper, color = group)) %>%
  fig_errorbar()
```

```{r}
huron <- dplyr::tibble(
  year = as.character(1875:1972), 
  level = as.vector(LakeHuron)
) %>% 
  dplyr::mutate(
    mn = level - 1,
    mx = level + 1
  )

g2(huron, asp(year, level, ymin = mn, ymax = mx)) %>%
  fig_line() %>%
  fig_ribbon() %>% 
  gauge_x_time(tickCount = 3)
```

```{r}
g2(cars, asp(speed, dist)) %>%
  fig_point() %>%
  fig_smooth()
```